---
title: "HW3"
author: "Zsiros, Gabriella"
date: "2023-02-21"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Target: 

predict fast growth of firms 
2010-2015

Structure:

# PART I: Probability prediction
## Probability prediction
Predict probabilities
## Cross-validation
Look at cross-validated performance and pick your favorite model
# PART II: Classification
## Loss function
Think about the business problem, and define your loss function 
(like FP=X dollars, FN=Y dollars)
## Classification threshold
For each model, predict probabilities, look for the optimal 
classification threshold, calculate expected loss with your loss function. 
## Average exprected loss
Pick the model that has the smallest average (over 5 folds) expected loss
# PART III Discussion of results
## Confusion table 
Show a confusion table (on a selected fold or holdout set)
## Evaluation
Discuss results, evaluate how useful your model may be

```{r}
#libraries
```


```{r}
#data read
df <- read.csv('https://osf.io/3qyut/download')
```


```{r}
summary(df)

#how many has full balance sheet 
table(df$balsheet_notfullyear)
```




financial variables are: cost of goods, amortization, current assets, 
current liabilities, exptra expenditure, extra income, extra profit loss, 
finished production, fixed assets, income before tax, intangible assets, 
inventories, liquid assets, material expenditure, net domestic sales, 
net export sales, personnel expenditure, profit/loss annual, total sales, 
shareholder equity, subscribed capital, tangible assets, wage bill

non numeric: balance sheet,  firm founded, exit year, number of CEO, 
share of foreign CEO, share of female CEO, avg birth year of CEO, 
average office time of CEOs, gender, CEO origin, HQ, HQ region,  (num) 


Possible growth influencing factors:

Total sales: Companies with increasing sales are more likely to experience fast growth.

Income before tax: A higher income before tax can indicate a company's ability to reinvest in its business and support growth.

Shareholder equity: Higher shareholder equity can indicate a company's ability to raise funds to support growth.

Tangible assets: Companies with higher tangible assets may have the resources to invest in growth opportunities.

Inventories: Increasing inventories can indicate that a company is preparing for future sales growth.

Personnel expenditure: Companies that invest in their employees may have a more skilled and motivated workforce that can support fast growth.
